---
title: Traffic Data Analysis
output: html_document
---

```{r, CACHE = TRUE}
library(reshape2)
library(ggplot2)
library(dplyr)
library(tidyr)
```
```{r, CACHE = TRUE}
mydata <- read.csv("all-semi-unique.csv")
nrow(mydata)
ncol(mydata)
names(mydata)
head(mydata)
summary(mydata)
```

Having a look at the internal structure and types of the columns.
```{r, CACHE = TRUE}
str(mydata)
```
All the columns that are of type 'Factor' and have only one level are going to be removed, they are not adding any new information to the data or creating patterns. Clean data by removing them.
Those columns are : ad.bgcl , ad.bgcls, ad.fncl, ad.fncls, ad.logo, ad.logo2x, ad.logoAndroidS, ad.logoAndroidH, ad.cm, ad.url, rd.cl.

```{r, CACHE = TRUE}
mydata <- subset.data.frame(mydata, select = -c(ad.bgcl , ad.bgcls, ad.fncl, ad.fncls, ad.logo, ad.logo2x, ad.logoAndroidS, ad.logoAndroidH, ad.cm, ad.url, rd.cl))
```
Check if there are still other columns that have no unique values. 
```{r, CACHE = TRUE}
col_ct = sapply(mydata, function(x) length(unique(x)))
length(col_ct[col_ct==1])
names(col_ct[col_ct==1])
```
Remove those columns.
```{r, CACHE = TRUE}
mydata = mydata[, !names(mydata) %in% names(col_ct[col_ct==1])]
```
Check for duplicate rows.
```{r, CACHE = TRUE}
nrow(mydata) - nrow(unique(mydata))
```
Checking ratio of NAs for every column
```{r, CACHE = TRUE}
sapply(mydata, function(x) sum(is.na(x))/(nrow(mydata)))
# barplot(sapply(mydata[,1:10], function(col) sum(is.na(col))),xlab = " columns ", ylab = " NA count ")
# barplot(sapply(mydata[,11:19], function(col) sum(is.na(col))),xlab = " columns ", ylab = " NA count ")
```
The values of the image columns have a high ratio of NAs, these images can either be of the user who posted the comment, or a picture they took of the street reported according to the website, do not find it useful, remove it to clear the data.

```{r, CACHE = TRUE}
mydata <- subset.data.frame(mydata, select = -c(rd.rp.rpImg, rd.rp.img, rd.img))
```
Break down the crawl date into date, time.
First set crawl date into a time-date context instead of a factor, extract from it the actual time of report using rd.rp.hr and rd.rp.mn which represent elapsed time since the report was posted in UTC, then add 2 hours to obtain cairo time. Then Get the day of week which corresponds to that date, finally get rid of those columns.
```{r, CACHE = TRUE}
mydata$crawl_date <-strptime(mydata$crawl_date, format = "%a %b %e %X UTC %Y", tz = "UTC")
str(mydata$crawl_date)
mydata$calculated_cairo_date <- as.POSIXct(mydata$crawl_date, tz = "UTC") - mydata$rd.rp.hr*60*60 - mydata$rd.rp.mn*60 + 2*60*60
mydata$Day.Of.Week <- weekdays(mydata$calculated_cairo_date)
mydata <- subset.data.frame(mydata, select = -c(crawl_date, rd.rp.hr, rd.rp.mn))
str(mydata)
```
Plotting numeric columns
```{r, CACHE = TRUE}
ncol(mydata)
numeric_data <- mydata[, sapply(mydata, is.numeric)]
d <- melt(numeric_data)
ggplot(d,aes(x = value)) + 
    facet_wrap(~variable,scales = "free") + 
    geom_histogram()
```
Checking uniqueness of each column value.
```{r}
uniqueness = sapply(mydata, function(x) length(unique(x)))
```
By taking a look at the html of the bey2ollak website, we can find out that 

Default Comment| Corresponding stid
---------------|--------------
7alawa:        |       1
lazeez:        |       2 
mashy:         |       3
za7ma:         |       4 
mafeesh 2amal: |       5


Extracting the hour of the day from the calculated time of the report.
Turning the hour of day and week day into factors with discrete levels.
Extracting data based on only on the five different road status IDs provided above.
```{r}
t.str <- strptime(mydata$calculated_cairo_date, "%Y-%m-%d %H:%M:%S")
mydata$Hour.Of.Day <- as.numeric(format(t.str, "%H"))
mydata$Day.Of.Week <- factor(mydata$Day.Of.Week)
mydata$Hour.Of.Day <- factor(mydata$Hour.Of.Day)
#mydata$rd.stid <- factor(mydata$rd.stid, ordered = T)
traffic_status <- subset(mydata, mydata$rd.stid %in% c(1,2,3,4,5))
```
Trying to plot the state of the traffic based on the mean of the provided status ids on the different days of the week.
```{r}
# ggplot(data = traffic_status, aes(x=Day.Of.Week)) + geom_histogram(aes(weights= density, fill=rd.stid)) + scale_fill_brewer(palette="Set3") + facet_wrap( ~ rd.stid, ncol=1)
Traffic_weekDay <- traffic_status %>% group_by(Day.Of.Week) %>% summarise(mean = mean(rd.stid))
Traffic_weekDay%>% ggplot(aes(x=Day.Of.Week, y=mean)) + geom_bar(stat="identity", width = 0.8)
```

Trying to plot the state of the traffic based on the mean of the provided status ids on the different hours of the day.
```{r}
Traffic_hour <- traffic_status %>% group_by(Hour.Of.Day) %>% summarise(mean = mean(rd.stid))
Traffic_hour%>% ggplot(aes(x=Hour.Of.Day, y=mean)) + geom_bar(stat="identity", width = 0.8)
```

Frequency of reporting days.
```{r, TIDY = TRUE, CACHE = TRUE}
qplot(mydata$Day.Of.Week, xlab = "Days of the week", ylab = "Reports count")
```
As can be seen from the histograms, the rd.rp.stid and rd.stid have a range of 10 discrete values.